---
title: "R Notebook"
output: html_notebook
---

setwd("/Users/zaynabsaeed/Documents/R_Assignments")

data_train <- read.csv("diabetic_data.csv", sep = ",", header = TRUE)
id_train <- read.csv("IDs_mapping.csv", sep = ",", header = TRUE)

install.packages("rmarkdown")
install.packages("plyr")
install.packages("ggplot2")

install.packages("reticulate")

library(plyr)
library(ggplot2)
library(rmarkdown)

install_miniconda(path = miniconda_path(), update = TRUE, force = FALSE)


```{r setup, include=FALSE, echo = TRUE}
library(reticulate)
use_python("/Users/zaynabsaeed/Library/r-miniconda/envs/r-reticulate/bin/python")
```




render("DataMiningProject1.Rmd")

#Phase 1

#saransh's code
#Prepare the data for analysis
#Plotting the bar graphs of various variables to see what values each variable has
```{r}
#Objective: to see if there is any bad data
ggplot(data_train, aes(gender)) + geom_bar(aes()) 
```

```{r}

#Outcome: gender has male, female and unknown/invalid which are all good
ggplot(data_train, aes(age)) + geom_bar(aes()) 
```

```{r}

#Outcome: no blank age ranges all data seems good
ggplot(data_train, aes(weight)) + geom_bar(aes()) 


#Outcome: weight has most of the data as ? so we need to replace it with NA
#Replacing ? with NA
data_train$weight <- ifelse (test = data_train$weight == '?', yes = NA, no = data_train$weight)
```


```{r}

ggplot(data_train, aes(race)) + geom_bar(aes())
#Outcome: race has some ? data so we need to replace it with NA
#Coverting ? to Unknown
data_train$race <- revalue(x=data_train$race,replace=c("?"="Unknown"))
```

```{r}

ggplot(data_train, aes(payer_code)) + geom_bar(aes()) 
#Outcome: Payer code has some ?'s
#Replacing ? with Unknown
data_train$payer_code <- revalue(x=data_train$payer_code,replace=c("?"= "Unknown"))
```

```{r}
ggplot(data_train, aes(medical_specialty)) + geom_bar(aes()) 
#Outcome: medical_speciality has some ?'s
#Replacing ? with Unknown
data_train$medical_specialty <- revalue(x=data_train$medical_specialty,replace=c("?"= "Unknown"))
```

```{r}

#table for seeing frequencies in diag_1
table(data_train$diag_1)
#table for seeing frequencies in diag_2
table(data_train$diag_2)
#table for seeing frequencies in diag_3
table(data_train$diag_3)
```

```{r}

#Outcome: all of the three have ?'s that will be replaced by Unknowns
data_train$diag_1 <- revalue(x=data_train$diag_1,replace=c("?"= "Unknown"))
data_train$diag_2 <- revalue(x=data_train$diag_2,replace=c("?"= "Unknown"))
data_train$diag_3 <- revalue(x=data_train$diag_3,replace=c("?"= "Unknown"))

```



```{python}
!pip install pandas
```



#Phase 2


```{python}
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt

diabetes_data = pd.read_csv('diabetic_data.csv')

# Extract the different values for the age attribute
ages = diabetes_data['age']
age_categories = np.unique(np.array(ages))

```

```{python}

# Create the new categories
# [0,30)
age_0 = age_categories[:3]
# [30,60)
age_1 = age_categories[3:6]
# [60,100]
age_2 = age_categories[6:]
new_ages = []
age_occurrences = [0, 0, 0]

print(f'First age category: {age_0}')
print(f'Second age category: {age_1}')
print(f'Third age category: {age_2}')

```

```{python}

# Create a new list replacing the old age intervals with the new ones
for age in ages.values:
    if age in age_0:
        new_ages.append(0)
        age_occurrences[0] += 1
    elif age in age_1:
        new_ages.append(1)
        age_occurrences[1] += 1
    elif age in age_2:
        new_ages.append(2)
        age_occurrences[2] += 1
    else:
        new_ages.append('NaN')

# Add the new category values to the DataFrame
new_age_series = pd.Series(new_ages)
diabetes_data['new_ages'] = new_age_series

new_age_category = np.unique(np.array(new_age_series))

print(f'New age categories: {new_age_category}')
print(f'Occurrences of each age category: {age_occurrences}')
```

```{python}

# Plot the occurrences of each category
plt.bar(new_age_category, age_occurrences)
plt.show()
```



```{r}

#age ANALYSIS

#converting age to numerical form
age.num <- revalue(x = data_train$age, replace = c("[0-10)" = 0, "[10-20)" = 1,
                                                   "[20-30)" = 2, "[30-40)" = 3,
                                                   "[40-50)" = 4, "[50-60)" = 5, 
                                                   "[60-70)" = 6, "[70-80)" = 7,
                                                   "[80-90)" = 8, "[90,100)" = 9))

data_train$age_numeric <-as.numeric(levels(age.num))[age.num]

hist(data_train$age_numeric,xlab="age_numeric", 
     main = "Historgram of age_numeric")

#Standardizing the data
data_train$age_numeric_z <- scale(x=data_train$age_numeric)

#Identify outliers
data_outliers_age <- data_train[which(data_train$age_numeric_z < -3 | 
                                                data_train$age_numeric_z > 3),]
#852 outliers detected
```


```{r}

#number_diagnoses ANALYSIS

#Standardizing the data
data_train$number_diagnoses_z <- scale(x=data_train$number_diagnoses)

#Identify outliers
data_outliers_number_diagnoses <- data_train[which(data_train$number_diagnoses_z < -3 | 
                                        data_train$number_diagnoses_z > 3),]
#281 outliers detected

#MEDICATION TYPES ANALYSIS

#metformin 
#converting metformin to numerical form
met.num <- revalue(x = data_train$metformin, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$metformin_numeric <-as.numeric(levels(met.num))[met.num]

hist(data_train$metformin_numeric,xlab="metformin_numeric", 
     main = "Historgram of metformin_numeric")

#Standardizing the data
data_train$metformin_numeric_z <- scale(x=data_train$metformin_numeric)

#Identify outliers
data_outliers_metformin <- data_train[which(data_train$metformin_numeric_z < -3 | 
                                        data_train$metformin_numeric_z > 3),]
#1067 outliers detected
```


```{r}

#repaglinide
#converting repaglinide to numerical form
rep.num <- revalue(x = data_train$repaglinide, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$repaglinide_numeric <-as.numeric(levels(rep.num))[rep.num]

hist(data_train$repaglinide_numeric,xlab="repaglinide_numeric", 
     main = "Historgram of repaglinide_numeric")

#Standardizing the data
data_train$repaglinide_numeric_z <- scale(x=data_train$repaglinide_numeric)

#Identify outliers
data_outliers_repaglinide <- data_train[which(data_train$repaglinide_numeric_z < -3 | 
                                              data_train$repaglinide_numeric_z > 3),]
#1539 outliers detected


```

```{r}

#nateglinide
#converting nateglinide to numerical form
nat.num <- revalue(x = data_train$nateglinide, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$nateglinide_numeric <-as.numeric(levels(nat.num))[nat.num]

hist(data_train$nateglinide_numeric,xlab="natglinide_numeric", 
     main = "Historgram of natglinide_numeric")

#Standardizing the data
data_train$natglinide_numeric_z <- scale(x=data_train$nateglinide_numeric)

#Identify outliers
data_outliers_natglinide <- data_train[which(data_train$natglinide_numeric_z < -3 | 
                                                data_train$natglinide_numeric_z > 3),]
#703 outliers detected

```


```{r}

#chlorpropamide
#converting chlorpropamide to numerical form
chl.num <- revalue(x = data_train$chlorpropamide, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$chlorpropamide_numeric <-as.numeric(levels(chl.num))[chl.num]

hist(data_train$chlorpropamide_numeric,xlab="chlorpropamide_numeric", 
     main = "Historgram of chlorpropamide_numeric")

#Standardizing the data
data_train$chlorpropamide_numeric_z <- scale(x=data_train$chlorpropamide_numeric)

#Identify outliers
data_outliers_chlorpropamide <- data_train[which(data_train$chlorpropamide_numeric_z < -3 | 
                                               data_train$chlorpropamide_numeric_z > 3),]
#86 outliers detected

```



```{r}

#glimepiride
#converting glimepiride to numerical form
gli.num <- revalue(x = data_train$glimepiride, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$glimepiride_numeric <-as.numeric(levels(gli.num))[gli.num]

hist(data_train$glimepiride_numeric,xlab="glimepiride_numeric", 
     main = "Historgram of glimepiride_numeric")

#Standardizing the data
data_train$glimepiride_numeric_z <- scale(x=data_train$glimepiride_numeric)

#Identify outliers
data_outliers_glimepiride <- data_train[which(data_train$glimepiride_numeric_z < -3 | 
                                                   data_train$glimepiride_numeric_z > 3),]
#4997 outliers detected

```



```{r}

#acetohexamide
#converting acetohexamide to numerical form
ace.num <- revalue(x = data_train$acetohexamide, replace = c("No" = 0, "Steady" = 1))
#suprisingly, "Down" and "Up" were not detected in this column

data_train$acetohexamide_numeric <-as.numeric(levels(ace.num))[ace.num]

hist(data_train$acetohexamide_numeric,xlab="acetohexamide_numeric", 
     main = "Historgram of acetohexamide_numeric")

#Standardizing the data
data_train$acetohexamide_numeric_z <- scale(x=data_train$acetohexamide_numeric)

#Identify outliers
data_outliers_acetohexamide <- data_train[which(data_train$acetohexamide_numeric_z < -3 | 
                                                data_train$acetohexamide_numeric_z > 3),]
#1 outlier detected

```



```{r}

#glipizide
#converting glipizide to numerical form
glip.num <- revalue(x = data_train$glipizide, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$glipizide_numeric <-as.numeric(levels(glip.num))[glip.num]

hist(data_train$glipizide_numeric,xlab="glipizidenumeric", 
     main = "Historgram of glipizide_numeric")

#Standardizing the data
data_train$glipizide_numeric_z <- scale(x=data_train$glipizide_numeric)

#Identify outliers
data_outliers_glipizide <- data_train[which(data_train$glipizide_numeric_z < -3 | 
                                                data_train$glipizide_numeric_z > 3),]
#770 outliers detected

```





```{r}

#glyburide
#converting glyburide to numerical form
gly.num <- revalue(x = data_train$glyburide, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$glyburide_numeric <-as.numeric(levels(gly.num))[gly.num]

hist(data_train$glyburide_numeric,xlab="glyburide_numeric", 
     main = "Historgram of glyburide_numeric")

#Standardizing the data
data_train$glyburide_numeric_z <- scale(x=data_train$glyburide_numeric)

#Identify outliers
data_outliers_glyburide <- data_train[which(data_train$glyburide_numeric_z < -3 | 
                                              data_train$glyburide_numeric_z > 3),]
#812 outliers detected

```


```{r}

#tolbutamide
#converting tolbutamide to numerical form
tol.num <- revalue(x = data_train$tolbutamide, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$tolbutamide_numeric <-as.numeric(levels(tol.num))[tol.num]

hist(data_train$tolbutamide_numeric,xlab="tolbutamide_numeric", 
     main = "Historgram of tolbutamide_numeric")

#Standardizing the data
data_train$tolbutamide_numeric_z <- scale(x=data_train$tolbutamide_numeric)

#Identify outliers
data_outliers_tolbutamide <- data_train[which(data_train$tolbutamide_numeric_z < -3 | 
                                              data_train$tolbutamide_numeric_z > 3),]
#23 outliers detected

```



```{r}

#pioglitazone
#converting tolbutamide to numerical form
pio.num <- revalue(x = data_train$pioglitazone, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$pioglitazone_numeric <-as.numeric(levels(pio.num))[pio.num]

hist(data_train$pioglitazone_numeric,xlab="pioglitazone_numeric", 
     main = "Historgram of pioglitazone_numeric")

#Standardizing the data
data_train$pioglitazone_numeric_z <- scale(x=data_train$pioglitazone_numeric)

#Identify outliers
data_outliers_pioglitazone <- data_train[which(data_train$pioglitazone_numeric_z < -3 | 
                                                data_train$pioglitazone_numeric_z > 3),]
#7210 outliers detected

```





```{r}

#rosiglitazone
#converting rosiglitazone to numerical form
ros.num <- revalue(x = data_train$rosiglitazone, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$rosiglitazone_numeric <-as.numeric(levels(ros.num))[ros.num]

hist(data_train$rosiglitazone_numeric,xlab="rosiglitazone_numeric", 
     main = "Historgram of rosiglitazone_numeric")

#Standardizing the data
data_train$rosiglitazone_numeric_z <- scale(x=data_train$rosiglitazone_numeric)

#Identify outliers
data_outliers_rosiglitazone <- data_train[which(data_train$rosiglitazone_numeric_z < -3 | 
                                                 data_train$rosiglitazone_numeric_z > 3),]
#6278 outliers detected

```





```{r}

#acarbose
#converting acarbose to numerical form
aca.num <- revalue(x = data_train$acarbose, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$acarbose_numeric <-as.numeric(levels(aca.num))[aca.num]

hist(data_train$acarbose_numeric,xlab="acarbose_numeric", 
     main = "Historgram of acarbose_numeric")

#Standardizing the data
data_train$acarbose_numeric_z <- scale(x=data_train$acarbose_numeric)

#Identify outliers
data_outliers_acarbose <- data_train[which(data_train$acarbose_numeric_z < -3 | 
                                                  data_train$acarbose_numeric_z > 3),]
#308 outliers detected

```



```{r}

#miglitol
#converting miglitol to numerical form
mig.num <- revalue(x = data_train$miglitol, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$miglitol_numeric <-as.numeric(levels(mig.num))[mig.num]

hist(data_train$miglitol_numeric,xlab="miglitol_numeric", 
     main = "Historgram of miglitol_numeric")

#Standardizing the data
data_train$miglitol_numeric_z <- scale(x=data_train$miglitol_numeric)

#Identify outliers
data_outliers_miglitol <- data_train[which(data_train$miglitol_numeric_z < -3 | 
                                             data_train$miglitol_numeric_z > 3),]
#38 outliers detected

```


```{r}

#troglitazone
#converting troglitazone to numerical form
tro.num <- revalue(x = data_train$troglitazone, replace = c("No" = 0, "Steady" = 1))
#suprisingly, "Down" and "Up" were not detected in this column

data_train$troglitazone_numeric <-as.numeric(levels(tro.num))[tro.num]

hist(data_train$troglitazone_numeric,xlab="troglitazone_numeric", 
     main = "Historgram of troglitazone_numeric")

#Standardizing the data
data_train$troglitazone_numeric_z <- scale(x=data_train$troglitazone_numeric)

#Identify outliers
data_outliers_troglitazone <- data_train[which(data_train$troglitazone_numeric_z < -3 | 
                                             data_train$troglitazone_numeric_z > 3),]
#3 outliers detected


```

```{r}

#tolazamide
#converting tolazamide to numerical form
tolz.num <- revalue(x = data_train$tolazamide, replace = c("No" = 0, "Steady" = 1, "Up" = 2))
#suprisingly, "Down" was not detected in this column

data_train$tolazamide_numeric <-as.numeric(levels(tolz.num))[tolz.num]

hist(data_train$tolazamide_numeric,xlab="tolazamide_numeric", 
     main = "Historgram of tolazamide_numeric")

#Standardizing the data
data_train$tolazamide_numeric_z <- scale(x=data_train$tolazamide_numeric)

#Identify outliers
data_outliers_tolazamide <- data_train[which(data_train$ttolazamide_numeric_z < -3 | 
                                                 data_train$tolazamide_numeric_z > 3),]
#0 outliers detected


```


```{r}

#examide
#converting examide to numerical form
exa.num <- revalue(x = data_train$examide, replace = c("No" = 0))
#"Down", "Steady", and "Up" were not present in column

data_train$examide_numeric <-as.numeric(levels(exa.num))[exa.num]

hist(data_train$examide_numeric,xlab="examide_numeric", 
     main = "Historgram of examide_numeric")

#Standardizing the data
data_train$examide_numeric_z <- scale(x=data_train$examide_numeric)

#Identify outliers
data_outliers_examide <- data_train[which(data_train$examide_numeric_z < -3 | 
                                             data_train$examide_numeric_z > 3),]
#0 outliers detected
```





```{r}

#citoglipton
#converting citoglipton to numerical form
cit.num <- revalue(x = data_train$citoglipton, replace = c("No" = 0))

data_train$citoglipton_numeric <-as.numeric(levels(cit.num))[cit.num]

hist(data_train$citoglipton_numeric,xlab="citoglipton_numeric", 
     main = "Historgram of citoglipton_numeric")

#Standardizing the data
data_train$citoglipton_numeric_z <- scale(x=data_train$citoglipton_numeric)

#Identify outliers
data_outliers_citoglipton <- data_train[which(data_train$citoglipton_numeric_z < -3 | 
                                             data_train$citoglipton_numeric_z > 3),]
#0 outliers detected

```


```{r}

#insulin
#converting insulin to numerical form
ins.num <- revalue(x = data_train$insulin, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$insulin_numeric <-as.numeric(levels(ins.num))[ins.num]

hist(data_train$insulin_numeric,xlab="insulin_numeric", 
     main = "Historgram of insulin_numeric")

#Standardizing the data
data_train$insulin_numeric_z <- scale(x=data_train$insulin_numeric)

#Identify outliers
data_outliers_insulin <- data_train[which(data_train$insulin_numeric_z < -3 | 
                                             data_train$insulin_numeric_z > 3),]
#0 outliers detected

```



```{r}

#glyburide.metformin
#converting glyburide.metformin to numerical form
glym.num <- revalue(x = data_train$glyburide.metformin, replace = c("No" = 0, "Down" = 1, "Steady" = 2, "Up" = 3))

data_train$glyburide.metformin_numeric <-as.numeric(levels(glym.num))[glym.num]

hist(data_train$glyburide.metformin_numeric,xlab="glyburide.metformin_numeric", 
     main = "Historgram of glyburide.metformin_numeric")

#Standardizing the data
data_train$glyburide.metformin_numeric_z <- scale(x=data_train$glyburide.metformin_numeric)

#Identify outliers
data_outliers_glyburide.metformin <- data_train[which(data_train$glyburide.metformin_numeric_z < -3 | 
                                            data_train$glyburide.metformin_numeric_z > 3),]
#706 outliers detected

```



```{r}


#glipizide.metformin
#converting glipizide.metformin to numerical form
glipm.num <- revalue(x = data_train$glipizide.metformin, replace = c("No" = 0, "Steady" = 1))
#"Down" and "Up" not present in column

data_train$glipizide.metformin_numeric <-as.numeric(levels(glipm.num))[glipm.num]

hist(data_train$glipizide.metformin_numeric,xlab="glipizide.metformin_numeric", 
     main = "Historgram of glipizide.metformin_numeric")

#Standardizing the data
data_train$glipizide.metformin_numeric_z <- scale(x=data_train$glipizide.metformin_numeric)

#Identify outliers
data_outliers_glipizide.metformin <- data_train[which(data_train$glipizide.metformin_numeric_z < -3 | 
                                                        data_train$glipizide.metformin_numeric_z > 3),]
#13 outliers detected

```



```{r}

#glimepiride.pioglitazone
#converting glimepiride.pioglitazone to numerical form
glip.num <- revalue(x = data_train$glimepiride.pioglitazone, replace = c("No" = 0, "Steady" = 1))
#"Down" and "Up" not present in column

data_train$glimepiride.pioglitazone_numeric <-as.numeric(levels(glip.num))[glip.num]

hist(data_train$glimepiride.pioglitazone_numeric,xlab="glimepiride.pioglitazone_numeric", 
     main = "Historgram of glimepiride.pioglitazone_numeric")

#Standardizing the data
data_train$glimepiride.pioglitazone_numeric_z <- scale(x=data_train$glimepiride.pioglitazone_numeric)

#Identify outliers
data_outliers_glimepiride.pioglitazone <- data_train[which(data_train$glimepiride.pioglitazone_numeric_z < -3 | 
                                                        data_train$glimepiride.pioglitazone_numeric_z > 3),]
#1 outlier detected


```


```{r}

#metformin.rosiglitazone
#converting metformin.rosiglitazone to numerical form
metr.num <- revalue(x = data_train$metformin.rosiglitazone, replace = c("No" = 0, "Steady" = 1))
#"Down" and "Up" not present in column

data_train$metformin.rosiglitazone_numeric <-as.numeric(levels(metr.num))[metr.num]

hist(data_train$metformin.rosiglitazone_numeric,xlab="metformin.rosiglitazone_numeric", 
     main = "Historgram of metformin.rosiglitazone_numeric")

#Standardizing the data
data_train$metformin.rosiglitazone_numeric_z <- scale(x=data_train$metformin.rosiglitazone_numeric)

#Identify outliers
data_outliers_metformin.rosiglitazone <- data_train[which(data_train$metformin.rosiglitazone_numeric_z < -3 | 
                                                        data_train$metformin.rosiglitazone_numeric_z > 3),]
#2 outliers detected
```




```{r}

#metformin.pioglitazone
#converting metformin.pioglitazone to numerical form
metp.num <- revalue(x = data_train$metformin.pioglitazone, replace = c("No" = 0, "Steady" = 1))
#"Down" and "Up" not present in column


data_train$metformin.pioglitazone_numeric <-as.numeric(levels(metp.num))[metp.num]

hist(data_train$metformin.pioglitazone_numeric,xlab="metformin.pioglitazone_numeric", 
     main = "Historgram of metformin.pioglitazone_numeric")

#Standardizing the data
data_train$metformin.pioglitazone_numeric_z <- scale(x=data_train$metformin.pioglitazone_numeric)

#Identify outliers
data_outliers_metformin.pioglitazone <- data_train[which(data_train$metformin.pioglitazone_numeric_z < -3 | 
                                                        data_train$metformin.pioglitazone_numeric_z > 3),]
#1 outlier detected
```




#Phase 3
```{r}

#Histograms
#hist(data_train$gender)
#hist(data_train$race)

hist(data_train$admission_type_id)
hist(data_train$num_lab_procedures)
hist(data_train$num_procedures)
hist(data_train$num_medications)
hist(data_train$time_in_hospital)
hist(data_train$number_diagnoses)

```

```{r}

#diabetesMed Analysis
dia_med_num <- revalue(x = data_train$diabetesMed, replace = c("Yes" = 1, "No" = 0))

data_train$diabetesMed_numeric <-as.numeric(levels(dia_med_num))[dia_med_num]

hist(data_train$diabetesMed_numeric,xlab="diabetesMed_numeric", 
     main = "Historgram of diabetesMed_numeric")


#ggplot
ggplot(data_train,aes(diabetesMed_numeric)) + geom_bar() + coord_flip()

#Standardizing the data
data_train$diabetesMed_numeric_z <- scale(x=data_train$diabetesMed_numeric)

#Identify outliers
data_outliers_diabetesMed <- data_train[which(data_train$diabetesMed_numeric_z < -3 | 
                                                data_train$diabetesMed_numeric_z > 3),]


#readmitted overlay
ggplot(data_train, aes(diabetesMed_numeric)) + geom_bar(aes(fill=readmitted)) + coord_flip()
#0 means they didn't recieve medication
#1 means they did recieve medication

#normalized
ggplot(data_train, aes(diabetesMed_numeric)) + geom_bar(aes(fill=readmitted), position = "fill") + coord_flip()

#contingency table
t.v1 <-table(data_train$readmitted,data_train$diabetesMed_numeric)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v2



```



```{r}

#admission_type_id analysis

#Standardizing the data
data_train$admission_type_id_z <- scale(x=data_train$admission_type_id)

#Identify outliers
data_outliers_admission_type_id <- data_train[which(data_train$admission_type_id_z < -3 | 
                                                      data_train$admission_type_id_z > 3),] 
#341 outliers detected

#readmitted overlay
ggplot(data_train, aes(admission_type_id)) + 
  geom_bar(aes(fill = readmitted)) + coord_flip() 

#normalized
ggplot(data_train, aes(admission_type_id)) + 
  geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

#contingency table
t.v1 <-table(data_train$readmitted,data_train$admission_type_id)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v2

```


```{r}

#time_in_hospital analysis

#Standardizing the data
data_train$time_in_hospital_z <- scale(x=data_train$time_in_hospital)

#Identify outliers
data_outliers_time_in_hospital <- data_train[which(data_train$time_in_hospital_z < -3 | 
                                                     data_train$time_in_hospital_z > 3),] 
#1042 outliers detected

#readmitted overlay
ggplot(data_train, aes(time_in_hospital)) + geom_bar(aes(fill = readmitted)) + coord_flip() 

#normalized
ggplot(data_train, aes(time_in_hospital)) + geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

#contingency table
t.v1 <-table(data_train$readmitted,data_train$time_in_hospital)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v2

```


```{r}

#num_lab_procedures analysis

#Standardizing the data
data_train$num_lab_procedures_z <- scale(x=data_train$num_lab_procedures)

#Identify outliers
data_outliers_num_lab_procedures <- data_train[which(data_train$num_lab_procedures_z < -3 | 
                                                       data_train$num_lab_procedures_z > 3),]
#43 Outliers detected


#readmitted overlay
ggplot(data_train, aes(num_lab_procedures)) + geom_bar(aes(fill = readmitted)) + coord_flip() 

#normalized
ggplot(data_train, aes(num_lab_procedures)) + geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

#contingency table
t.v1 <-table(data_train$readmitted,data_train$num_lab_procedures)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v2

```


```{r}

#num_medications analysis

#Standardizing the data
data_train$num_medications_z <- scale(x=data_train$num_medications)

#Identify outliers
data_outliers_num_medications <- data_train[which(data_train$num_medications_z < -3 | 
                                                    data_train$num_medications_z > 3),]
#1361 outliers detected

#readmitted overlay
ggplot(data_train, aes(num_medications)) + geom_bar(aes(fill = readmitted)) + coord_flip() 

#normalized
ggplot(data_train, aes(num_medications)) + geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

#contingency table
t.v1 <-table(data_train$readmitted,data_train$num_medications)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v2

```


```{r}

#number_diagnosis analysis

#Standardizing the data
data_train$number_diagnoses_z <- scale(x=data_train$number_diagnoses)

#Identify outliers
data_outliers_number_diagnoses<- data_train[which(data_train$number_diagnoses_z < -3 | 
                                                    data_train$number_diagnoses_z > 3),]
#0 outliers detected

#readmitted overlay
ggplot(data_train, aes(number_diagnoses)) + geom_bar(aes(fill = readmitted)) + coord_flip() 

#normalized
ggplot(data_train, aes(number_diagnoses)) + geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

#contingency table
t.v1 <-table(data_train$readmitted,data_train$number_diagnoses)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v2

```

```{r}

#gender analysis

#readmitted overlay
ggplot(data_train, aes(gender)) + geom_bar(aes(fill = readmitted)) + coord_flip() 

#normalized
ggplot(data_train, aes(gender)) + geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

#contingency table
t.v1 <-table(data_train$readmitted,data_train$gender)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v1.rnd <- round(prop.table(t.v1,margin=2)*100,1)

t.v1
t.v2

```

```{r}

#race analysis

#readmitted overlay 
ggplot(data_train, aes(race)) + geom_bar(aes(fill = readmitted)) + coord_flip()

#normalized
ggplot(data_train, aes(race)) + geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

t.v1 <-table(data_train$readmitted,data_train$race)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v1.rnd <- round(prop.table(t.v1,margin=2)*100,1)

t.v1
t.v2
```


```{python}

import pandas as pd
import numpy as np
import matplotlib.pyplot as plt

diabetes_data = pd.read_csv('diabetic_data.csv')
```


```{python}

possible_diagnosis = [
                        "circulatory",
                        "respiratory",
                        "digestive",
                        "diabetes",
                        "injury",
                        "musculoskeletal",
                        "genitourinary",
                        "neoplasms",
                        "other"
                     ]


def determine_diag_map(icd9_code):
    """
    This function determines the diagnostic map
    depending on the table above. It Assigns a
    category referred by the group name in lower case
    :param icd9_code: String
    """
    if icd9_code.find("E") != -1 or \
            icd9_code.find("V") != -1:
        return "other"
    elif icd9_code.find("?") != -1:
        return None
    else:
        icd9_code = float(icd9_code)
        if 390 <= icd9_code <= 459 or \
                icd9_code == 785:
            return "circulatory"
        elif 460 <= icd9_code <= 519 or \
                icd9_code == 786:
            return "respiratory"
        elif 520 <= icd9_code <= 579 or \
                icd9_code == 787:
            return "digestive"
        elif 250 <= icd9_code < 251 or \
                icd9_code == 787:
            return "diabetes"
        elif 800 <= icd9_code <= 999:
            return "injury"
        elif 710 <= icd9_code <= 739:
            return "musculoskeletal"
        elif 580 <= icd9_code <= 629 or \
                icd9_code == 788:
            return "genitourinary"
        elif 140 <= icd9_code <= 239:
            return "neoplasms"
        else:
            return "other"


#
for diagnosis_type in possible_diagnosis:
    diabetes_data[diagnosis_type] = ""
# Bin the diagnoses
diabetes_data["diag_1_binned"] = list(map(determine_diag_map, diabetes_data["diag_1"].values))
diabetes_data["diag_2_binned"] = list(map(determine_diag_map, diabetes_data["diag_2"].values))
diabetes_data["diag_3_binned"] = list(map(determine_diag_map, diabetes_data["diag_3"].values))

print(f"Post-bin shape: {diabetes_data.tail()}")
# Create a new data frame using the binned diagnoses
diagnosis_df = diabetes_data[["diag_1_binned",
                              "diag_2_binned",
                              "diag_3_binned"]].stack().str.get_dummies().sum(level=0)

diagnosis_df[diagnosis_df > 1] = 1

print(diagnosis_df.tail())


```

