---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

data_train <- read.csv("diabetic_data.csv", sep = ",", header = TRUE)
id_train <- read.csv("IDs_mapping.csv", sep = ",", header = TRUE)

install.packages("rmarkdown")
library(rmarkdown)
install.packages("plyr")
install.packages("ggplot2")
library(plyr)
library(ggplot2)
setwd("/Users/zaynabsaeed/Documents/R_Assignments")
render("DataMiningProjectNotebook.Rmd")



#target is readmitted





#Histograms
#hist(data_train$gender)
#hist(data_train$race)
```{r message=TRUE}
hist(data_train$admission_type_id)
hist(data_train$num_lab_procedures)
hist(data_train$num_procedures)
hist(data_train$num_medications)
hist(data_train$time_in_hospital)
hist(data_train$number_diagnoses)
```



```{r}

#diabetesMed Analysis
dia_med_num <- revalue(x = data_train$diabetesMed, replace = c("Yes" = 1, "No" = 0))

data_train$diabetesMed_numeric <-as.numeric(levels(dia_med_num))[dia_med_num]

hist(data_train$diabetesMed_numeric,xlab="diabetesMed_numeric", 
     main = "Historgram of diabetesMed_numeric")

#ggplot
ggplot(data_train,aes(diabetesMed_numeric)) + geom_bar() + coord_flip()

#Standardizing the data
data_train$diabetesMed_numeric_z <- scale(x=data_train$diabetesMed_numeric)

#Identify outliers
data_outliers_diabetesMed <- data_train[which(data_train$diabetesMed_numeric_z < -3 | 
                                    data_train$diabetesMed_numeric_z > 3),]

#readmitted overlay
ggplot(data_train, aes(diabetesMed_numeric)) + geom_bar(aes(fill=readmitted)) + coord_flip()
#0 means they didn't recieve medication
#1 means they did recieve medication

#normalized
ggplot(data_train, aes(diabetesMed_numeric)) + geom_bar(aes(fill=readmitted), position = "fill") + coord_flip()

#contingency table
t.v1 <-table(data_train$readmitted,data_train$diabetesMed_numeric)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v2
```




```{r echo=FALSE}
#admission_type_id analysis

#Standardizing the data
data_train$admission_type_id_z <- scale(x=data_train$admission_type_id)

#Identify outliers
data_outliers_admission_type_id <- data_train[which(data_train$admission_type_id_z < -3 | 
                                                data_train$admission_type_id_z > 3),] 
#341 outliers detected

#readmitted overlay
ggplot(data_train, aes(admission_type_id)) + 
  geom_bar(aes(fill = readmitted)) + coord_flip() 

#normalized
ggplot(data_train, aes(admission_type_id)) + 
  geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

#contingency table
t.v1 <-table(data_train$readmitted,data_train$admission_type_id)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v2

```


```{r}

#time_in_hospital analysis

#Standardizing the data
data_train$time_in_hospital_z <- scale(x=data_train$time_in_hospital)

#Identify outliers
data_outliers_time_in_hospital <- data_train[which(data_train$time_in_hospital_z < -3 | 
                                                      data_train$time_in_hospital_z > 3),] 
#1042 outliers detected

#readmitted overlay
ggplot(data_train, aes(time_in_hospital)) + geom_bar(aes(fill = readmitted)) + coord_flip() 

#normalized
ggplot(data_train, aes(time_in_hospital)) + geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

#contingency table
t.v1 <-table(data_train$readmitted,data_train$time_in_hospital)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v2

```

```{r}

#num_lab_procedures analysis

#Standardizing the data
data_train$num_lab_procedures_z <- scale(x=data_train$num_lab_procedures)

#Identify outliers
data_outliers_num_lab_procedures <- data_train[which(data_train$num_lab_procedures_z < -3 | 
                                                data_train$num_lab_procedures_z > 3),]
#43 Outliers detected


#readmitted overlay
ggplot(data_train, aes(num_lab_procedures)) + geom_bar(aes(fill = readmitted)) + coord_flip() 

#normalized
ggplot(data_train, aes(num_lab_procedures)) + geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

#contingency table
t.v1 <-table(data_train$readmitted,data_train$num_lab_procedures)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v2


```

```{r}

#num_medications analysis

#Standardizing the data
data_train$num_medications_z <- scale(x=data_train$num_medications)

#Identify outliers
data_outliers_num_medications <- data_train[which(data_train$num_medications_z < -3 | 
                                                       data_train$num_medications_z > 3),]
#1361 outliers detected

#readmitted overlay
ggplot(data_train, aes(num_medications)) + geom_bar(aes(fill = readmitted)) + coord_flip() 

#normalized
ggplot(data_train, aes(num_medications)) + geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

#contingency table
t.v1 <-table(data_train$readmitted,data_train$num_medications)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v2
```

```{r}


#number_diagnosis analysis

#Standardizing the data
data_train$number_diagnoses_z <- scale(x=data_train$number_diagnoses)

#Identify outliers
data_outliers_number_diagnoses<- data_train[which(data_train$number_diagnoses_z < -3 | 
                                                       data_train$number_diagnoses_z > 3),]
#0 outliers detected

#readmitted overlay
ggplot(data_train, aes(number_diagnoses)) + geom_bar(aes(fill = readmitted)) + coord_flip() 

#normalized
ggplot(data_train, aes(number_diagnoses)) + geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

#contingency table
t.v1 <-table(data_train$readmitted,data_train$number_diagnoses)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v2

```


```{r}
#gender analysis

#readmitted overlay
ggplot(data_train, aes(gender)) + geom_bar(aes(fill = readmitted)) + coord_flip() 

#normalized
ggplot(data_train, aes(gender)) + geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

#contingency table
t.v1 <-table(data_train$readmitted,data_train$gender)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v1.rnd <- round(prop.table(t.v1,margin=2)*100,1)

t.v1
t.v2
```

```{r}

#race analysis

#readmitted overlay 
ggplot(data_train, aes(race)) + geom_bar(aes(fill = readmitted)) + coord_flip()

#normalized
ggplot(data_train, aes(race)) + geom_bar(aes(fill = readmitted), position = "fill") + coord_flip() 

t.v1 <-table(data_train$readmitted,data_train$race)

t.v2 <-addmargins(A=t.v1,FUN=list(total=sum),quiet=TRUE)

t.v1.rnd <- round(prop.table(t.v1,margin=2)*100,1)

t.v1
t.v2
```
